---
title: "Plotting-Players"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Plotting-Players}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Installation procedure

Since the fpl is a package hosted on Github, you will need to download
it with the "remotes" package.

```{r setup}
# Package set-up
# install.packages("remotes")
# library(remotes)
# remotes::install_github("maxwell-miller-geo/fpl")
library(fpl)
```

# Download the data

from the official Premier League API at
'<https://fantasy.premierleague.com/api/bootstrap-static/>' using the
function get_premier_stats()

```{r get-data}
# Download Premier League Data
premier_data <- get_premier_stats() # Grabs current Premier League Data

# Filter data by players
players <- filter_players(premier_data, get_photos = F) # use default photos on
# Sample of player data
head(players[c("team", "web_name", "form", "points_per_game", "cost", "position",  "goals_scored")])
```

The output player data is a data.frame and can be analyzed or selected
with additional packages or in-built R scripts. Below is an example of
finding the top 10 scoring players for a given position.

```{r, dataframe}
# Select position
forwards <- players[players$position == "FWD",] # selecting using data.frame methods

# Order the forwards
column_order <- order(forwards$goals_scored, decreasing = TRUE)
# Create organize data.frame

forwards_ordered <- forwards[column_order,]
# Goals Scored
head(forwards_ordered[1:6]) # Select first 6 cols to display
```
## Filtering using data.table
```{r, datatable}
# Another way to order columns with an additional package: data.table
library(data.table)

# Add data.table functionality with data.table
players_dt <- data.table(players)
# order by points per game, selected first 6 columns for display
ordered_forwards <- players_dt[position == "FWD", ][order(-points_per_game)]
# Points Per Game
head(ordered_forwards[,1:6]) # Select first 6 cols to display - data.table way
```

# Now let's plot!

The primary things to remember before plotting is to look at the names
of the columns that are present.

```{r, colsPresent}
head(colnames(players), 15)
```

# Choose two columns you want to plot

For this example, we will plot "cost" versus the "total_points",
grouping by forwards, and displaying the players by name. Note: There is
some "smart" matching when inputting the columns, but it works best if
you can input exactly the column names. It will take a few seconds
depending on the group.

Some plots may have overlapping points, if there are to many players! 

```{r, plotting1,fig.width=9, fig.height=6, out.width="100%"}
fpl_plot(players, x = "cost", y = "total_points", group = "names", pos = "FWD")
```

## Plotting by Team Icon
```{r, plotting2,fig.width=9, fig.height=6, out.width="100%"}
#fpl_plot(players, x = "cost", y = "total_points", group = "teams", pos = "FWD")
```

## Plotting by Team Photo
Note: Some photos might be missing
```{r, plotting3, fig.width=9, fig.height=6, out.width="100%"}
#fpl_plot(players, x = "cost", y = "total_points", group = "photos", pos = "FWD")
```

## To many players, no problem
The plotting function works for almost any subset of the players data.frame.
So, you can subset or remove some players from a table and then plot! This is
especially the case with midfielders. Top percent filters by total_points.
```{r, plotting4, fig.width=9, fig.height=6, out.width="100%"}
fpl_plot(players, "cost", y = "form", group = "names", pos = "MID", top_percent = .85) # take top 85 percent
```

